Proposed Solution

### 2017-01-25 CAV17/secs/abs.tex

System failures resulting from configuration errors 
are major reasons for compromised reliability of today's software
systems. Although a lot of program verification techniques 
have been proposed, verifying configuration files is still 
an open problem,
because 1) software configurations are typically written in
poorly structured and untyped ``languages'', and 
2) specifying rules for configuration 
verification is challenging in practice.

>>>>

System failures resulting from configuration errors 
are major reasons for compromised reliability of today's software
systems. Although a lot of program verification techniques 
have been proposed, verifying configuration files is, nevertheless, 
a challenging problem,
because 1) software configurations are typically written in
poorly structured and untyped ``languages'', and 
2) specifying rules for configuration 
verification is challenging in practice.

### 2017-01-25 CAV17/secs/intro.tex

@@

Configuration errors are one of the most important root causes of today's software system failures~\cite{xu15systems, yin11anempirical}.
In their empirical study, Yin {\em et al.}~\cite{yin11anempirical} report that about 31\% of system failures were caused by misconfiguration problems and only 20\% were caused by bugs in program code. 
Misconfigurations, result in various system-wide problems, such as security vulnerabilities, application crashes, severe disruptions in software functionality, and incorrect program executions~\cite{zhang14encore, yuan11context, xu13do, xu15hey}.

While many efforts have been proposed to check, troubleshoot, diagnose, and repair configuration errors~\cite{attariyan10automating,su07autobash, whitaker04configuration},
  those tools mainly post-factum and cannot give the safety guarantees available from powerful static analysis tools in program verification~\cite{Leino10Dafny, PiskacWZ14, BobotFMP15}.
Existing techniques for program verification do not fit this application domain, as configuration files often lack specifications and structure.
There are rarely formal, or any, specifications on the constraints of the key values.
In the case of MySQL, the canonical specification is a 5240 page reference manual~\cite{mysql-manual15} to describe valid configurations and behavior.
We aim to automatically learn a formal specification the encodes this information by examining a set of example configuration files.

In previous work~\cite{santolucitoCAV}, we used a set of manually verified files by an domain expert order to learn specifications.
This approach is expensive and still prone to error by the expert.
Related work on configuration verification has used machine learning

While the grammar of configuration files tends to be simple (a series of key value pairs with some additional context), the keys and values themselves lack often structure and are untyped plain text.

   1) lack of a specification which would describe properties of configuration files, and 
   2) a program structure of configuration files -- they are mainly a sequence of entries assigning some value to system variables. 
The language in which configuration files are written does not adhere to a specific grammar or syntax. 
In particular, the entries in configuration files are untyped. 
Moreover, there are surprisingly few rules specifying constraints on entries and there is no explicit structure policy for the entries.
Thus, automated verification of configuration files would be highly desirable~\cite{wang04automatic, zhang14encore, xu15systems}.

To overcome these obstacles, researchers proposed approaches based on
statistical analysis and learning%
~\cite{wang04automatic, zhang14encore, yuan11context}
that try to infer rules and policies about 
how configuration files are constructed.
Rather than explicitly specifying entries' types or rules, 
these efforts focused on learning policies from a sample dataset.
The learned rules are usually constructed in the following way: 
for every entry in a configuration file, 
they check if it deviates from a ``typical'' value, \ie, 
a value computed from a large training set 
consisting of configuration files.
If the entry is significantly different from a typical value, 
they suspect that it could be a potential configuration error.

The downside of these is that their learning approaches 
are limited to simplistic configuration errors, 
such as type errors and syntax errors, or they 
heavily rely on template-based inference~\cite{zhang14encore}. 
%As we mentioned earlier, 
%because configuration files are not written in a typed language 
%with a fixed syntax,
%type errors and syntax errors are relatively rare and easy to detect. 
Many sophisticated configuration errors, 
which are difficult to template in practice, 
cannot be detected in this way. 
As an illustration of such an error, consider entry ordering errors, 
which happens when some entries are inserted in a wrong order. 
For example, if in a PHP configuration file 
an entry {\tt extension = mysql.so} appears 
before {\tt extension = recode.so},  
it would lead to a crash error, where the Apache server cannot start 
due to the segmentation fault error. The correct ordering 
should be {\tt extension = recode.so} before 
{\tt extension = mysql.so}~\cite{yin11anempirical}.
These types of errors cannot be detected by existing
learning efforts, 
since it is hard to build a corresponding template~\cite{xu15systems}. 
%we would need to capture ordering relations of all the entries. 
%Some of them can appear in an arbitrary ordering, 
%and the resulting templates would be too large 
%and not efficient for a practical use.

>>>>

Configuration errors are one of the most important root causes of today's 
software system failures~\cite{xu15systems, yin11anempirical}.
In their empirical study, Yin {\em et al.}~\cite{yin11anempirical} report 
that about 31\% of system failures were caused by misconfiguration problems
and only 20\% were caused by bugs in program code. 
Misconfigurations, in practice, may result in various system-wide problems,
including security vulnerabilities, application crashes, severe disruptions
in software functionality, 
and incorrect program executions~\cite{xu15systems, xu13do, xu15hey}.  
Therefore, how to automatically verify configuration files is an open
problem~\cite{xu15systems}.

To the best of our knowledge, there have been several prior 
efforts that attempt to verify 
configuration files~\cite{santolucitoCAV, xu16early,
zhang14encore, huang15confvalley}.
However, state-of-the-art efforts either are impractical to be used
in reality, or can only deal with simplistic configuration errors.
In general, these efforts fall into two categories.

\squishlist

\item On the one hand are tools that can detect sophisticated 
configuration errors, \eg, ConfigC~\cite{santolucitoCAV}. 
However, these tools heavily rely on datasets containing 100\% 
correct configuration files to extract configuration rules.
Existing investigation studies~\cite{wang04automatic, yin11anempirical}
have demonstrated determining or obtaining 100\% correct configuration
files to drive rules is almost impossible in reality. 
Without 100\% correct datasets, these tools would, unfortunately, 
work poorly.

\item On the other hand, practical efforts, 
\eg, EnCore~\cite{zhang14encore} and
ConfValley~\cite{huang15confvalley}, can only 
detect simplistic configuration errors (\eg, value range errors 
and simple integer correlation errors), but cannot detect
sophisticated configuration errors. 
Consider entry ordering errors, 
which happen when some entries are inserted in a wrong order
in the configuration files. 
For example, in a PHP configuration file, 
an entry \code{extension=mysql.so} appears 
before \code{extension=recode.so},  
it would cause Apache server to crash. The correct ordering 
should be \code{extension=recode.so} before 
\code{extension=mysql.so}~\cite{yin11anempirical}.
Nevertheless, these types of errors cannot be detected by above existing
efforts.

\squishend

The goal of this work is to propose and build a {\em practical} 
configuration
verification tool capable of checking {\em sophisticated} 
configuration errors (more details
for sophisticated configuration errors in Sec.~\ref{sec:motiv}).
Achieving this goal requires us addressing two main obstacles:
1) the lack of a specification describing properties of configuration
files, and 2) the structure of configuration files---they are 
mainly a sequence of entries assigning some values to system
parameters (called keywords).
In other words, the ``language'' in which configuration 
files are written does not adhere to any specific grammar or syntax. 
Moreover, the entries in configuration files 
lack type or rules specifying constraints on entries, 
and there is no explicit structure policy for these entries.

This paper presents \app, the first automatic verification framework
for general software configurations, by overcoming the above two
obstacles. In general, \app verifies a given configuration file $F$ 
through three steps:

\para{Step~1: Translator}
\app takes as input a large training set of configuration files
belonging the same system as $F$,
and we do not have any assumption on the correctness of these files.
\app analyzes the training set, and generates a well-structured and
probabilistically-typed intermediate representations.

\para{Step~2: Learning}
\app derives configuration rules by running {\em rule association
learning} algorithm to learn the generated intermediate representations.
The insight of our design is: every file in the training set
contains several different errors and these errors only appear 
in a small percentage of files. Based on this insight,
we are able, with the help of the probabilistic cutoff, to learn a
set of configuration rules. These rules, in general, specify 
which properties the entries in configuration files need to satisfy. 
One can see this learning process as a way of deriving  
a specification for configuration files.  
Once a specification is obtained, we can do verification that 
uses these learned rules to check the correctness of 
the configuration files
of interest and detect potential errors.

\para{Step~3: Rule graph analysis}
To ensure the learned rules are good enough,
\app employs a further analysis stage to refine the learned rules.
In general, \app builds a graph to model the learned rules,
and analyzes this graph's properties, ranking the importance
and relevance of these rules. Finally, \app outputs
a set of refined rules as the eventual specification
to check the target configuration file $F$.

We have implemented a \app prototype, and evaluated it based on
real-world datasets. In our verification experiments, 
we first use a public configuration dataset~\cite{configdataset}, which
contains 261 incorrect MySQL configuration files, as our
training set. We then use \app to learn rules from this training set
and refine these rules. Finally, we check sophisticated 
errors of real configuration files crawled from github,
and detect many errors in these widely-used configuration files.

In summary, we make the following contributions.

\com{

@@

%Therefore, automated verification of configuration 
%files would be highly
%desirable~\cite{wang04automatic, zhang14encore, xu15systems}.

### 2017-01-25 CAV17/secs/postanal.tex
comment out implementation

\app can output the set of learned rules from the training set as
described in $\S$\ref{sec-learn}, recalling that a rule is 
an implication relationship of the form $r = X \implies p(X,Y)$.
This data is sufficient for \app to perform its core task of
configuration file validation, but can be used for more than 
simple rule checking. By re-interpreting
rules as being the edges of a {\bf Rule Graph}, we use tools from network
theory to extract information
about the configuration space. We inspect properties of the Rule Graph 
to triage reported errors by those most likely to be valid,
and even to estimate the complexity of configuration settings.

Accessibility of the Rule Graph is a hallmark of the association rule
learning technique applied by \app \cite{someOtherGraphStuff}.
In contrast, the model output from a neural network is not as easily 
amenable to analysis \cite{nnAnalysis1, nnAnalysis2}, making it harder
to build on the work of previously developed methods and tools.

The following section provides a precise definition of the Rule Graph
and the metrics derived for the purposes of ranking reported errors
and complexity analysis.

@@

We define the \textit{Rule Graph} as a directed hypergraph $H = (V,E)$,
   with vertices $V = \{ keywords \}$ and labeled, weighted edges $E = \{ (V_s, V_t, l, w) \}$.
The set of edges is constructed from the learned rules, using the keyword sets as source and targets, the predicates as labels, and the confidence as weights:

\begin{align*}
\forall r \in & Learn(\trainingSet).\ \exists e\ \in E.\\
              & V_s = X_r \land V_t=Y_r \land l=p_r \land w = confidence(r)
\end{align*}

@@

    $$E_{V_1, V_2} = \{ \left( V_s, V_t, l, w \right) \in E \ \mid \ \exists v_1 \in V_1, V_s \land \exists v_2 \in V_2, V_t \}$$

Sometimes we will write a single vertex $v$ in our subscript, as
notational convenience for the singleton set containing $v$.

@@

    $$|E_{v_1, v_2}| = \sum_{(X, Y, l, w) \in E_{v_1, v_2}} w$$

@@

\para {Definition}
We define a measure of degree $\mathcal{D}(v)$ for each vertex $v$ as the sum of in-degree
and out-degree defined in the usual way. Explicitly, for a vertex $x \in V$:

@@

The more rules of high confidence are extracted for a keyword by the learner, 
the higher the $DEGREE$ of the corresponding vertex in the Rule Graph.
By classifying keywords in this way, we may extract a simple way to
order the {\it errors} reported by \app by estimated importance.

We hypothesize that keywords (specifically their corresponding vertices)
of low $DEGREE$ will be rarer configuration
parameters where rules learned are more likely to be governed by 
tehchnical necessity, rather than industry convention. As such, errors
reported involving low-degree keywords are more likely to be errors
of high significance and should be presented with high importance
to users of \app.

Specifically, for an error reported by \app on a rule $r$ involving
keywords $K$, we may rank the errors simply by:

    $$RANK(r) = avg( DEGREE(K) )$$

@@

We may also use the Rule Graph to advance our general knowledge
of the configuration space, outside the strict confines of a
verification system. As an example, we briefly present a heuristic
for configuration file complexity based on the topology of the
Rule Graph. This measure of complexity could be used by software
organizations to manage configuration files in much the same way
as Kolomogrov complexity~\cite{kolomogrov} is used to manage code - identifying
potentially brittle configurations for targeted refactoring.

@@

For a configuration file with a set of keywords $K$ for a system
and a Rule Graph $H = (V, E)$, we define our complexity measure:

@@

The complexity measure, $\mathcal{C}$, can be thought of as an
extension of the na\"ive line-counting measure of complexity.
When a keyword in the configuration file is present in the Rule Graph,
we may consider the set $E_{k, K}$ to be all learned rules involving
keyword $k$ that
are {\it relevant} to the configuration file being examined. the set
$E_{k, V}$ denotes {\it all} learned rules involving $k$.
Given these sets, we may think of $\frac{|E_{k, K}|}{|E_{k, V}|}$
as representing the amount that $k$ is constrained in the current
configuration file relative to how much it could be constrained
in the global configuration space. The more constrained a configuration
keyword in a particular configuration file, the {\it less} it should
contribute to the complexity (hence $1 * (1 - \frac{|E_{k, K}|}{|E_{k, V}|})$).

If a keyword is not constrained at all in the current configuration
file or is not present in the Rule Graph, we revert to the standard
counting metric of complexity.

>>>>

\app can output the set of learned rules from the training set as
described in $\S$\ref{sec-learn}. Recall that a rule is 
an implication relationship of the form $r = X \implies p(X,Y)$.
This data is sufficient for \app to perform its core task of
configuration file validation, but can be used for more than 
simple rule checking. By interpreting the rules as a graph (which
we call the {\bf Rule Graph}), we can 
use tools from graph theory to extract information
about the configuration space. We inspect properties of this 
Rule Graph to sort reported errors by those most likely to be valid.
To demonstrate the additional value of the Rule Graph, we
also use it to estimate the complexity of configuration settings.

Accessibility of the Rule Graph is a useful property of the association 
rule learning technique applied by \app.
In contrast, the model output from a neural network is not as easily 
amenable to analysis \cite{nnAnalysis1, nnAnalysis2}, making it harder
to build on the work of previously developed methods and tools.
The following section provides a precise definition of the Rule Graph
and the metrics derived for the purposes of ranking reported errors
and complexity analysis.

@@

We define the \textit{Rule Graph} as a directed hypergraph $H = (V,E)$,
   with vertices $V = \{ keywords \}$ and labeled, weighted edges $E = \{ (V_s, V_t, l, w) \}$.
The set of edges is constructed from the learned rules, using the source and target keyword sets as sources and targets respectively, the predicates as labels, and the confidence as weights:
%
\begin{align*}
\forall r \in & Learn(\trainingSet).\ \exists e\ \in E.\\
              & V_s = X_r \land V_t=Y_r \land l=p_r \land w = confidence(r)
\end{align*}

@@

\begin{align*}
    E_{V_1, V_2} = \{ \left( V_s, V_t, l, w \right) \in E \ \mid \ & \exists v_1 \in V_1 \land v_1 \in V_s \\
    & \land \exists v_2 \in V_2 \land v_2 \in V_t \}
\end{align*}

We can think of $E_{V_1, V_2}$ as being the subset of edges in $E$ 
such that each source set $V_s$ shares a vertex with $V_1$
and each target set $V_t$ shares a vertex with $V_2$.
We denote a naked vertex $v$ in our subscripts as notational 
convenience for the singleton set containing that vertex $v$.

@@

    $$|E| = \sum_{(X, Y, l, w) \in E} w$$

@@

We define a measure of degree $\mathcal{D}(v)$ for each vertex $v$ as the sum of 
in-degree and out-degree. Explicitly, for a vertex $x \in V$:

@@

We may now use this measure to rank our errors.
The more rules of high confidence are extracted for a keyword by the learner, 
the higher the $\mathcal{D}(v)$ of the corresponding vertex in the Rule Graph.
In our final analysis, we use this classification to
order the reported {\it errors} by estimated importance.

Keywords (specifically their corresponding vertices)
of low $\mathcal{D}(v)$ may be rarer configuration
parameters where rules learned are more likely to be governed by 
tehchnical necessity, rather than industry convention. As such, errors
reported involving low-degree keywords are more likely to be errors
of high significance and should be presented with high importance
to users of \app.

Specifically, for an error reported by \app on a rule $r$ involving
keywords $K$, we rank the errors by:

    $$RANK(r) = \frac{\sum_{k \in K} \mathcal{D}(k)}{|K|}$$

@@

We may also use the Rule Graph to advance our general knowledge
of the configuration space, outside the strict confines of a
verification system. As an example, we present a heuristic
for configuration file complexity based on the topology of the
Rule Graph. This measure of complexity could be used by software
organizations to manage configuration files in much the same way
as Kolomogrov complexity~\cite{kolomogrov} is used to manage code - identifying
potentially brittle configurations for targeted refactoring.

@@

For a configuration file with a set of keywords $K$
and a Rule Graph $H = (V, E)$, we define our complexity measure:

@@

The complexity measure, $\mathcal{C}$, can be thought of as an
extension of the na\"ive line-counting measure of complexity.
When a keyword in the configuration file is present in the Rule Graph,
we may consider the set $E_{k, K}$ to be all learned rules involving
keyword $k$ that
are {\it relevant} to the configuration file being examined. 
The set $E_{k, V}$ denotes {\it all} learned rules involving $k$.
Given these sets, we may think of $\frac{|E_{k, K}|}{|E_{k, V}|}$
as representing the amount that $k$ is constrained in the current
configuration file relative to how much it could be constrained
in the global configuration space. The more constrained a configuration
keyword in a particular configuration file, the {\it less} it should
contribute to the complexity (hence $1 * (1 - \frac{|E_{k, K}|}{|E_{k, V}|})$).
If a keyword is not constrained at all in the current configuration
file or is not present in the Rule Graph, we revert to the standard
counting metric of complexity.

While an in-depth evaluation of the complexity metric presented here
is out of scope for this paper, we use this measure to
demonstrate the flexibility of the rule graph, and potential
for further applications.
